---
title: "RNAseq"
author: "Da"
date: '`r format(Sys.time(), "%a %b %d %Y - %X")`'
output: html_document
---

RNA-seq data analysis
=======================

This project is an analysis of RNAseq data on mouse Gcn1 and Atf4 knockdown cells and WT control cells.


Data source
-----------

The raw data for this project comes from Jonanthan Hernandez Lab at NIH

- RawData/filteredCounts has the filtered gene counts generated using [CCBR Pipeliner](https://ccbr.github.io/Pipeliner/). Genes with >= 1 CPM in >= 2 samples are included.
- RawData/result_webScrap.txt is gene Entrez IDs scrapped from NCBI using /analysis/spath_forDa.py
- RawData/filteredCounts/sampletable.txt also has metadata including sampleName, fileName, condition and label.


Objectives
--------

- Generate differential expression using Limma Voom.
- Perform over-representation analysis(ORA) and Gene Set Enrichment Analysis (GSEA) using Clusterprofiler.
- Investigate the genes and pathways regulated by Gcn1 and Atf4 either independently or together (overlap).


Analysis code is written for R version 3.6.1

---

Summary
--------
Sinha and Hernandez et al. identified the general control of amino acid synthesis 1-like 1 gene (Gcn1) in an in vivo genetic screen to isolate drivers of metastatic outgrowth. Preliminary in vitro and in vivo experiments suggest that Gcn1 drives tumor growth and metastasis during glutamine deprivation by activating the master regulator of the integrated stress response pathway ATF4. 

This dataset has 6 samples: two control sample (SCR) and four Gcn1 knockdown samples (SH1 and SH2) under 0.5 mM treatment. First the knock-down of Gcn1 is verified by importing bigwig/bam files into IGV and checking the exons of Gcn1 gene. SH1 and SH2 has lower counts on the first few exons. RSEM counts were normalized using the voom algorithm from the Limma R package (version 3.40.6). These normalized counts were used for clustering and data visualization. PCA shows SH1 knockdown samples have larger variation compared to SCR controls than SH2 compared to SCR. Heatmap generated using the top 500 most variable genes across samples visually confirmed this. 

The Limma package was used to test for differential gene expression between experimental conditions. Significant differentially expressed genes were identified with a false-discovery rate ≤ 0.05 and a log(fold change) > 1. The number of DEGs in SH1_0pt5mM-SCR_0pt5mM: 3300; SH2_0pt5mM-SCR_0pt5mM: 1028; SCR_4mM-SCR_0pt5mM: 244. Since SCR_4mM-SCR_0pt5mM serves as negative control, the DEGs in this contrast that overlap with two other contrasts were excluded for subsequent analysis. 

Over Representation Analysis (ORA) is used to identify biological functions or processes are over-represented (enriched) in differentially expressed genes (DEGs).Pathway enrichment was performed using pre-ranked Gene Set Enrichment Analysis (GSEA) with the KEGG and Reactome genesets from the Molecular Signatures Database.


```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.width=14, fig.height=9)
```

```{r load library}
suppressMessages(library(rgl))
suppressMessages(library(knitr))
suppressMessages(library('edgeR'))
suppressMessages(library('statmod'))
suppressMessages(library(gplots))
suppressMessages(library('reshape'))
suppressMessages(library('limma'))
suppressMessages(library('geneplotter'))
suppressMessages(library(filesstrings))
suppressMessages(library(ff))
suppressMessages(library(Glimma))
suppressMessages(library(dplyr))
suppressMessages(library(RColorBrewer))
suppressMessages(library(pheatmap))
suppressMessages(library(clusterProfiler))
suppressMessages(library(argparse))
suppressMessages(library(org.Hs.eg.db))
suppressMessages(library(org.Mm.eg.db))
suppressMessages(library(ReactomePA))
suppressMessages(library(ggplot2))
suppressMessages(library(EnhancedVolcano))
suppressMessages(library(ggfortify))
suppressMessages(library(ggrepel))
suppressMessages(library(VennDiagram))
suppressMessages(library(biomaRt))
suppressMessages(library(enrichplot))
suppressMessages(library(gridExtra))
suppressMessages(library(gplots))

```


```{r set working, result directories}
setwd("~/Desktop/active_projects/ccbr1016-RNAseq/analysis/")
analysis_dir = "~/Desktop/active_projects/ccbr1016-RNAseq/analysis/"
results_dir = "~/Desktop/active_projects/ccbr1016-RNAseq/analysis/results/"
DEGs_dir = "~/Desktop/active_projects/ccbr1016-RNAseq/analysis/processedData/DEGs/"
normalized_data_dir = "~/Desktop/active_projects/ccbr1016-RNAseq/analysis/processedData/normalized/"
```


```{r use limma voom to compute differential expression}
limmaDEG = function(DIR,FILE1,FILE2, CONTRASTS){
  
  sampleinfo=read.delim(FILE1)
  x = read.delim(FILE2,row.names=1)
  colnames(x)=as.character(sampleinfo[,4])
  
  myconditions = as.factor(sampleinfo$condition)
  dge_before = DGEList(counts=x,group=myconditions)
  dge_before <- calcNormFactors(dge_before,method="TMM")
  
  Group <- factor(sampleinfo$condition)
  design=model.matrix(~0+Group)
  contras=unlist(strsplit(CONTRASTS, split=" "))  
  cons=c()
  for(i in seq(1, length(contras), by = 2))
  {{
    cons=c(cons,paste(contras[i],"-",contras[i+1],sep=""))
  }}
  
  v1 <- voom(as.matrix(x),design,plot=F,normalize="quantile")
  
  cat("voom normalization\n")
  
  sf = v1$E/log2((x/colSums(x))*1000000)
  rn=rownames(v1$E)
  ensID=apply(array(as.character(rn)),1,function(z) unlist(strsplit(z, "\\|"))[1])
  gene=apply(array(as.character(rn)),1,function(z) unlist(strsplit(z, "\\|"))[2])
  mydata=cbind(ensID,gene,v1$E)
  
  sf=cbind(ensID,gene,sf)
  nb=length(contras)/2
  colnames(design) <- levels(Group)
  fit <- lmFit(v1,design)
  contrast.matrix <- makeContrasts(contrasts=cons,levels=design)
  fitb <- contrasts.fit(fit, contrast.matrix)
  ebayes.fit=eBayes(fitb)
  dt = decideTests(ebayes.fit)
  # require(Glimma)
  glMDPlot(ebayes.fit, status=dt, counts=v1, groups=Group, side.main="Symbols")
  cat("lmFit calculated\n")
  
  write.table(v1$E,file=paste("limma_Voom_normalized_",cons[i],"_data.txt",sep=""),
              row.names=TRUE,col.names=NA,sep="\t",quote=FALSE)

  all.genes.con = topTable(ebayes.fit, coef = i, number=nrow(ebayes.fit), sort.by="none")
  
  all.genes.con$FC <- ifelse(all.genes.con$logFC<0, -1/(2^all.genes.con$logFC), 2^all.genes.con$logFC)
  final=all.genes.con
  
  x=rownames(all.genes.con)
  ensID=apply(array(as.character(x)),1,function(z) unlist(strsplit(z, "\\|"))[1])
  gene=apply(array(as.character(x)),1,function(z) unlist(strsplit(z, "\\|"))[2])
  all.genes.con=cbind(ensID,gene,all.genes.con)
  write.table(all.genes.con,file=paste("limma_DEG_",cons[i],"_all_genes.txt",sep=""),
              sep="\t",col.names=NA)
  
  cat("limma DEG table generated\n")
  
  #Reformatted DEG File
  limmaout=read.table(paste("limma_DEG_",cons[i],"_all_genes.txt",sep=""), header = TRUE)
  limmaout=limmaout[,which(names(limmaout) %in% c("X","gene","logFC","P.Value"))]
  limmaout$fc=2^limmaout$logFC
  down_reg=limmaout$logFC<0
  limmaout$fc[down_reg]=-1/limmaout$fc[down_reg]
  limmaout=limmaout[,c("X","gene","fc","logFC","P.Value")]
  colnames(limmaout)=c("ensid_gene","gene","fc","log2fc","pvalue")
  limmaout$fdr=p.adjust(limmaout$pvalue,method='fdr',n=length(limmaout$pvalue))
  limmaout$gsea_ranking_score=-log10(limmaout$pvalue)*sign(limmaout$log2fc)
  write.table(limmaout,file=paste("limma_DEG_",cons[i],"_all_genes.txt",sep=""),row.names=FALSE,col.names=TRUE,quote=FALSE,sep="\t")
  
  cat("Reformatted DEG table generated\n")
  file.rename(from = list.files(pattern = "glimma-plots"), to = paste0("glimma-plots_",cons[i],sep=""))
  path1 = paste0(analysis_dir, paste("glimma-plots_",cons[i],sep=""))
  path2 = paste0(results_dir, paste("glimma-plots_",cons[i],sep=""))
  file.move(path1, path2)
  file.move(list.files(pattern = "DEG"), DEGs_dir)
  file.move(list.files(pattern = "limma_Voom_normalized"), normalized_data_dir)
}
```

```{r generate DEGs}
limmaDEG(DIR=analysis_dir,
         FILE1="~/Desktop/active_projects/ccbr1016-RNAseq/rawData/filterdCounts/DEG_SH1_0pt5mM-SCR_0pt5mM_1_2/sampletable.txt",
         FILE2="~/Desktop/active_projects/ccbr1016-RNAseq/rawData/filterdCounts/DEG_SH1_0pt5mM-SCR_0pt5mM_1_2/RawCountFile_RSEM_genes_filtered.txt",
         CONTRASTS="SH1_0pt5mM SCR_0pt5mM")

limmaDEG(DIR=analysis_dir,
         FILE1="~/Desktop/active_projects/ccbr1016-RNAseq/rawData/filterdCounts/DEG_SH2_0pt5mM-SCR_0pt5mM_1_2/sampletable.txt",
         FILE2="~/Desktop/active_projects/ccbr1016-RNAseq/rawData/filterdCounts/DEG_SH2_0pt5mM-SCR_0pt5mM_1_2/RawCountFile_RSEM_genes_filtered.txt",
         CONTRASTS="SH2_0pt5mM SCR_0pt5mM")

limmaDEG(DIR=analysis_dir,
         FILE1="~/Desktop/active_projects/ccbr1016-RNAseq/rawData/filterdCounts/DEG_SCR_4mM-SCR_0pt5mM_1_2/sampletable.txt",
         FILE2="~/Desktop/active_projects/ccbr1016-RNAseq/rawData/filterdCounts/DEG_SCR_4mM-SCR_0pt5mM_1_2/RawCountFile_RSEM_genes_filtered.txt",
         CONTRASTS="SCR_4mM SCR_0pt5mM")

limmaDEG(DIR=analysis_dir,
         FILE1="~/Desktop/active_projects/ccbr1016-RNAseq/rawData/filterdCounts/DEG_experiment-control_1_2/sampletable.txt",
         FILE2="~/Desktop/active_projects/ccbr1016-RNAseq/rawData/filterdCounts/DEG_experiment-control_1_2/RawCountFile_RSEM_genes_filtered.txt",
         CONTRASTS="experiment control")

# remove the experiment-control DEG table from subsequent pathway analysis
toRemove = dir(path = DEGs_dir,pattern = "control")
file.remove(paste(DEGs_dir,toRemove, sep = "/"))

```


```{r include=FALSE}
#' Truncate string vector of ggplot axis label
#'
#' @param label    a ordered string vector
#' @param maxLen   max length of character (nchar) to show in label
#' @param maxWord  max count of words allowed to show in label
#' @param pattern  Word separater
#' @param dot      If true, three dots will added to truncated label
#'
#' @return a vector of truncated strings
#' @export
#'
#' @examples
short_label <- function(label, maxLen = 50, maxWord = 10, pattern = " ", dot = TRUE){
  l <- strsplit(label, pattern)
  short_label <- vector("character",length(l))
  
  for (i in seq_along(l)){
    truncated <- FALSE
    s <- l[[i]]
    if (length(s) > maxWord){
      ll <- paste(s[1:maxWord], collapse = " ")
      truncated <- TRUE
    }
    else{
      ll <- paste(s, collapse = " ")
    }
    
    if (nchar(ll) > maxLen){
      ll <- substr(ll, 1, maxLen)
      truncated <- TRUE
    }
    
    if (dot & truncated) ll <- paste(ll, "...",sep = " ")
    
    short_label[[i]] <- ll
  }
  attr(short_label, "pos") <- attr(label,"pos")
  return(short_label)
}
```

read in the matrices into a list and modify the names
```{r read data}
rawData_dir = "~/Desktop/active_projects/ccbr1016-RNAseq/rawData/"
geneinfo = read.table("~/Desktop/active_projects/ccbr1016-RNAseq/rawData/geneinfo.bed")
geneinfo$V5 = gsub('\\..*','', geneinfo$V5)
DEGs_dir = "~/Desktop/active_projects/ccbr1016-RNAseq/analysis/processedData/DEGs/"
my_files = list.files(path = DEGs_dir)
my_degs <- list()
for (i in my_files){
  print(i)
  my_degs[[i]] <- read.table(paste(DEGs_dir,i,sep = "/"),header = TRUE,sep='\t')
  my_degs[[i]]$ensid_gene <- gsub('\\..*','', my_degs[[i]]$ensid_gene)
}
names(my_degs) <- stringr::str_replace(my_files, pattern = "_all_genes.txt", replacement = "") %>% 
  stringr::str_replace(pattern = "limma_DEG_", replacement = "")

```

rank genes based on gsea_ranking_score, for GO gsea
```{r calculate degs; rank genes based on gsea_ranking_score, message=FALSE, warning=FALSE}
my_sig_degs = lapply(my_degs, function(x) subset(x, fdr < 0.05 & abs(fc) > 2))

my_degs_rankingScoresList <- lapply(my_degs, function(x){
  scores <- x$gsea_ranking_score
  names(scores) <- as.character(x$ensid_gene)
  scores <- sort(scores, decreasing = T)
})

```

Mapping DEGs to entrez IDs using ensemble, then symbol, then NCBI web scrapped IDs.  
```{r Converts EnsemblIDs to UNIPROTID; rank genes based on gsea_ranking_score, message=FALSE, warning=FALSE}
Entrez_from_webScrap <- read.table("~/Desktop/active_projects/ccbr1016-RNAseq/rawData/result_webScrap.txt", sep = '\t', stringsAsFactors = F, col.names = c("ensid_gene", "ENTREZID"))
Entrez_from_webScrap$ensid_gene <- gsub('.*=','', Entrez_from_webScrap$ensid_gene) 
Entrez_from_webScrap$ENTREZID = as.character(Entrez_from_webScrap$ENTREZID)

my_genes = data.frame()
for (x in my_degs) {
  temp_genes = x[, c("ensid_gene", "gene")]
  my_genes = dplyr::bind_rows(my_genes, temp_genes)
  my_genes = my_genes[!(duplicated(my_genes$ensid_gene)),]
}

Entrez_from_ensembl <- bitr(my_genes$ensid_gene, fromType = "ENSEMBL", toType = "ENTREZID", OrgDb = org.Mm.eg.db)
colnames(Entrez_from_ensembl)[colnames(Entrez_from_ensembl)=="ENSEMBL"] <- "ensid_gene"
unmapped_ensid <- my_genes[!(my_genes$ensid_gene%in%Entrez_from_ensembl$ensid_gene),] 
unmapped_ensid <- unmapped_ensid[!duplicated(unmapped_ensid$gene),]
ensid_symbol.data <- bitr(unique(unmapped_ensid$gene), fromType = "SYMBOL", toType = "ENTREZID", OrgDb = org.Mm.eg.db)
ensid_symbol.data <- ensid_symbol.data[!duplicated(ensid_symbol.data$SYMBOL),]
ensid_symbol_unmapped_ensid.data <- unmapped_ensid[(unmapped_ensid$gene%in%ensid_symbol.data$SYMBOL),]
Entrez_from_symbol = cbind(ensid_symbol_unmapped_ensid.data, ensid_symbol.data)
Entrez_from_symbol = Entrez_from_symbol[c("ensid_gene", "ENTREZID")]
Entrez_all <- dplyr::bind_rows(Entrez_from_ensembl, Entrez_from_symbol, Entrez_from_webScrap)
Entrez_unmapped = setdiff(my_genes$ensid_gene, Entrez_all$ensid_gene)

paste("Out of all", length(my_genes$ensid_gene), "DEGs:", length(Entrez_from_ensembl$ENTREZID), "were mapped to Entrez using Ensemble ID,", length(Entrez_from_symbol$ENTREZID), "mapped using gene symbol,",
length(Entrez_from_webScrap$ENTREZID), "mapped by NCBI web scraping.", length(setdiff(my_genes$ensid_gene, Entrez_all$ensid_gene)), "cannot be mapped.")
```
```{r unmapped genes, eval=FALSE, include=FALSE}
Entrez_unmapped_geneinfo<- geneinfo[(geneinfo$V5%in%Entrez_unmapped),]
paste("Umapped genes include",sum(Entrez_unmapped_geneinfo$V6=="TEC"),"TEC,",sum(Entrez_unmapped_geneinfo$V6=="lincRNA"),"lincRNA,",sum(Entrez_unmapped_geneinfo$V6=="processed_pseudogene"), "processed_pseudogene,", sum(Entrez_unmapped_geneinfo$V6=="antisense"),"antisense,", sum(Entrez_unmapped_geneinfo$V6=="protein_coding"), "protein_coding, etc")
```

rank genes based on gsea_ranking_score, for KEGG & Reactome gsea 
```{r ranking Entrez genes}
my_degs_entrez = lapply(my_degs, function(x){
  result <- merge(x, Entrez_all, by="ensid_gene") #adding a entrez column
})

my_sig_degs_entrez = lapply(my_degs_entrez, function(x) subset(x, fdr < 0.05 & abs(fc) > 2))

my_degs_entrez_rankingScoresList <- lapply(my_degs_entrez, function(x){
  scores <- x$gsea_ranking_score
  names(scores) <- as.character(x$ENTREZID)
  scores <- sort(scores, decreasing = T)
})
```

#### PCA
```{r fig.width=8, fig.height=7}
normalized_data_dir = "~/Desktop/active_projects/ccbr1016-RNAseq/analysis/processedData/normalized/"
normalizedData = list.files(normalized_data_dir, pattern = "control")
normalized_data = read.table(paste(normalized_data_dir, normalizedData, sep = "/"), header = T, sep = '\t', row.names = 1)
normalized_data$gene = gsub(".*[|]","", rownames(normalized_data))

#normalized_data = subset(normalized_data, select = -c(X,gene))
SH1_SH2_SCR_0.5 = subset(normalized_data, select=c(SCR_0pt5mM_A,SCR_0pt5mM_B,SH1_0pt5mM_A,
                                                   SH1_0pt5mM_B,SH2_0pt5mM_A,SH2_0pt5mM_B))

pca.SH1_SH2_SCR_0.5 = prcomp(t(SH1_SH2_SCR_0.5))
autoplot(pca.SH1_SH2_SCR_0.5,label=T,repel = TRUE, 
         label.size = 3,label.repel=T)+ theme_light()
```

#### heatmap for Top 500 most variable genes across samples
```{r}
var_genes  = apply(SH1_SH2_SCR_0.5, 1, var)
select_var <- names(sort(var_genes, decreasing=TRUE))[1:500]
select_var_count = SH1_SH2_SCR_0.5[select_var,]

mypalette <- brewer.pal(11,"RdYlBu")
morecols <- colorRampPalette(mypalette)

heatmap.2(data.matrix(select_var_count),col=rev(morecols(50)),trace="none", 
          main="Top 500 most variable genes across samples",scale="row",srtCol=20,cexCol=0.8,cexRow = 0.1)
```

#### volcano plot for differentially expressed genes
```{r}
myVolcanos = lapply(my_degs, function(x){
  EnhancedVolcano(x, 
                lab = x$gene,x = "log2fc", y = "fdr", ylab = bquote(~-Log[10]~FDR),FCcutoff = 1, 
                ylim= c(0, max(-log10(x$fdr), na.rm=TRUE) + 1),pCutoff = 0.05,
                legendPosition = 'top',titleLabSize = 22, subtitleLabSize = 20,
                subtitle = "Differential expression",captionLabSize= 18,
                transcriptPointSize = 0.8,transcriptLabSize = 4.5,legendLabSize = 18,
                legendIconSize = 4,legend = c("NS","log2FC > |1|","FDR < 0.05",
                           "significant (FDR < 0.05) & log2FC > |1|")) 
})

myVolcanos[[1]] + ggtitle(paste(names(myVolcanos[1])))
myVolcanos[[2]] + ggtitle(paste(names(myVolcanos[2])))
myVolcanos[[3]] + ggtitle(paste(names(myVolcanos[3])))
```
```{r message=FALSE, warning=FALSE, include=FALSE}
pdf(paste(results_dir, "volcano_", names(myVolcanos[1]), ".pdf", sep = ""), width = 11, height = 10)
myVolcanos[[1]] + ggtitle(paste(names(myVolcanos[1])))
dev.off()
pdf(paste(results_dir,"volcano_", names(myVolcanos[2]), ".pdf", sep = ""), width = 11, height = 10)
myVolcanos[[2]] + ggtitle(paste(names(myVolcanos[2])))
dev.off()
pdf(paste(results_dir,"volcano_", names(myVolcanos[3]), ".pdf", sep = ""), width = 11, height = 10)
myVolcanos[[3]] + ggtitle(paste(names(myVolcanos[3])))
dev.off()
```

#### number of differentially expressed genes for each contrast
```{r echo=FALSE}
paste("Number of DEGS in",names(my_sig_degs[1]), ":", length(my_sig_degs[[1]]$gene), sep=" ")
paste("Number of DEGS in",names(my_sig_degs[2]), ":", length(my_sig_degs[[2]]$gene), sep=" ")
paste("Number of DEGS in",names(my_sig_degs[3]), ":", length(my_sig_degs[[3]]$gene), sep=" ")

```

#### Three 0.5mM contrasts
```{r fig.align='center', fig.height=6, fig.width=12}
venn.plot_3Con0.5 <- venn.diagram(list(my_sig_degs[[1]]$gene,
                                     my_sig_degs[[2]]$gene,
                                     my_sig_degs[[3]]$gene), NULL, 
                                fill=c("blue", "red","grey"), lwd = 1, cat.cex = 1,
                                alpha=c(0.5,0.5,0.5), cex = 2, cat.fontface=4, 
                                category.names=c(paste(names(my_sig_degs[1])),
                                                 paste(names(my_sig_degs[2])),
                                                 paste(names(my_sig_degs[3]))))
grid.draw(venn.plot_3Con0.5)

```

#### exlude the DEGs in SCR_4mM-SCR_0pt5mM from other contrasts
```{r exlude control degs}
my_sig_degs[["SH2_0pt5mM-SCR_0pt5mM"]] = my_sig_degs[["SH2_0pt5mM-SCR_0pt5mM"]][(my_sig_degs[["SH2_0pt5mM-SCR_0pt5mM"]]$ensid_gene%in%
      setdiff(my_sig_degs[["SH2_0pt5mM-SCR_0pt5mM"]]$ensid_gene, my_sig_degs[["SCR_4mM-SCR_0pt5mM"]]$ensid_gene)),]

my_sig_degs[["SH1_0pt5mM-SCR_0pt5mM"]] = my_sig_degs[["SH1_0pt5mM-SCR_0pt5mM"]][(my_sig_degs[["SH1_0pt5mM-SCR_0pt5mM"]]$ensid_gene%in%
      setdiff(my_sig_degs[["SH1_0pt5mM-SCR_0pt5mM"]]$ensid_gene, my_sig_degs[["SCR_4mM-SCR_0pt5mM"]]$ensid_gene)),]

paste("Number of DEGS after filtering in ","SH1_0pt5mM-SCR_0pt5mM: ", length(my_sig_degs[["SH1_0pt5mM-SCR_0pt5mM"]]$gene), sep="")
paste("Number of DEGS after filtering in ","SH2_0pt5mM-SCR_0pt5mM: ", length(my_sig_degs[["SH2_0pt5mM-SCR_0pt5mM"]]$gene), sep="")
paste("Number of DEGS in ","SCR_4mM-SCR_0pt5mM: ", length(my_sig_degs[["SCR_4mM-SCR_0pt5mM"]]$gene), sep="")
```

### GO over-representation test (ORA)

#### cellular component ORA
```{r over-representation ORA, message=FALSE, warning=FALSE}
myCC_ORAs = Map(function(x,y) enrichGO(gene = as.character(x$ensid_gene), universe = as.character(y$ensid_gene), OrgDb = org.Mm.eg.db,keyType='ENSEMBL', ont = "CC", pAdjustMethod = "BH", pvalueCutoff = 0.01, qvalueCutoff = 0.05,readable = TRUE), x=my_sig_degs, y=my_degs)
paste("ff","dd",sep = "_")
paste(names(myCC_ORAs[1]))

merge_result(list(SH1_SCR_0.5mM = myCC_ORAs[["SH1_0pt5mM-SCR_0pt5mM"]], SH2_SCR_0.5mM = myCC_ORAs[["SH2_0pt5mM-SCR_0pt5mM"]],SCR_4mM_SCR_0.5mM = myCC_ORAs[["SCR_4mM-SCR_0pt5mM"]])) %>% 
  dotplot(., showCategory=5) + ggtitle("biological process")+ scale_y_discrete(label=short_label)

```


```{r include=FALSE}
path_out = "~/Desktop/active_projects/ccbr1016-RNAseq/analysis/processedData/ORA_GSEA/"
writeFile = function(data, name){
  fileName = paste(path_out, name, sep='')
  write.csv(data, fileName)
}

writeFile(myCC_ORAs[[1]], paste0("myCC_ORAs_",names(myCC_ORAs[1]),".csv"))
writeFile(myCC_ORAs[[2]], paste0("myCC_ORAs_",names(myCC_ORAs[2]),".csv"))
writeFile(myCC_ORAs[[3]], paste0("myCC_ORAs_",names(myCC_ORAs[3]),".csv"))
```


#### biological process ORA
```{r biological process ORA, message=FALSE, warning=FALSE}
myBP_ORAs = Map(function(x,y) enrichGO(gene = as.character(x$ensid_gene), universe = as.character(y$ensid_gene), OrgDb = org.Mm.eg.db,keyType='ENSEMBL', ont = "BP", pAdjustMethod = "BH", pvalueCutoff = 0.01, qvalueCutoff = 0.05,readable = TRUE), x=my_sig_degs, y=my_degs)

merge_result(list(SH1_SCR_0.5mM = myBP_ORAs[["SH1_0pt5mM-SCR_0pt5mM"]], SH2_SCR_0.5mM = myBP_ORAs[["SH2_0pt5mM-SCR_0pt5mM"]],SCR_4mM_SCR_0.5mM = myBP_ORAs[["SCR_4mM-SCR_0pt5mM"]])) %>% 
  dotplot(., showCategory=5) + ggtitle("biological process")+ scale_y_discrete(label=short_label)

```

```{r include=FALSE}
writeFile(myBP_ORAs[[1]], paste0("myCC_ORAs_",names(myBP_ORAs[1]),".csv"))
writeFile(myBP_ORAs[[2]], paste0("myCC_ORAs_",names(myBP_ORAs[2]),".csv"))
writeFile(myBP_ORAs[[3]], paste0("myCC_ORAs_",names(myBP_ORAs[3]),".csv"))
```


#### molecular function ORA
```{r molecular function ORA, message=FALSE, warning=FALSE}
myMF_ORAs = Map(function(x,y) enrichGO(gene = as.character(x$ensid_gene), universe = as.character(y$ensid_gene), OrgDb = org.Mm.eg.db,keyType='ENSEMBL', ont = "MF", pAdjustMethod = "BH", pvalueCutoff = 0.01, qvalueCutoff = 0.05,readable = TRUE), x=my_sig_degs, y=my_degs)

merge_result(list(SH1_SCR_0.5mM = myMF_ORAs[["SH1_0pt5mM-SCR_0pt5mM"]], SH2_SCR_0.5mM = myMF_ORAs[["SH2_0pt5mM-SCR_0pt5mM"]],SCR_4mM_SCR_0.5mM = myMF_ORAs[["SCR_4mM-SCR_0pt5mM"]])) %>% 
  dotplot(., showCategory=6) + ggtitle("molecular function")+ scale_y_discrete(label=short_label)
```

```{r include=FALSE}
writeFile(myMF_ORAs[[1]], paste0("myCC_ORAs_",names(myMF_ORAs[1]),".csv"))
writeFile(myMF_ORAs[[2]], paste0("myCC_ORAs_",names(myMF_ORAs[2]),".csv"))
writeFile(myMF_ORAs[[3]], paste0("myCC_ORAs_",names(myMF_ORAs[3]),".csv"))
```

#### KEGG ORA
```{r KEGG ORA, message=FALSE, warning=FALSE}
myKEGG_ORAs = Map(function(x,y) enrichKEGG(gene = as.character(x$ENTREZID), universe = as.character(y$ENTREZID), organism = 'mmu',
pvalueCutoff = 0.01, pAdjustMethod = "BH", qvalueCutoff = 0.05,minGSSize = 10, maxGSSize = 500), x=my_sig_degs_entrez, y=my_degs_entrez)

merge_result(list(SH1_SCR_0.5mM = myKEGG_ORAs[["SH1_0pt5mM-SCR_0pt5mM"]], SH2_SCR_0.5mM = myKEGG_ORAs[["SH2_0pt5mM-SCR_0pt5mM"]],SCR_4mM_SCR_0.5mM = myKEGG_ORAs[["SCR_4mM-SCR_0pt5mM"]])) %>% 
  dotplot(., showCategory=5) + ggtitle("KEGG ORA")+ scale_y_discrete(label=short_label)
```

```{r include=FALSE}
writeFile(myKEGG_ORAs[[1]], paste0("myKEGG_ORAs_",names(myMF_ORAs[1]),".csv"))
writeFile(myKEGG_ORAs[[2]], paste0("myKEGG_ORAs_",names(myMF_ORAs[2]),".csv"))
writeFile(myKEGG_ORAs[[3]], paste0("myKEGG_ORAs_",names(myMF_ORAs[3]),".csv"))

```


#### Reactome ORA
```{r Reactome ORA, message=FALSE, warning=FALSE}
myReactome_ORAs = Map(function(x,y) enrichPathway(gene = as.character(x$ENTREZID), universe = as.character(y$ENTREZID), organism = 'mouse', pvalueCutoff = 0.01, pAdjustMethod = "BH", qvalueCutoff = 0.05, minGSSize = 10, maxGSSize = 500), x=my_sig_degs_entrez, y=my_degs_entrez)

merge_result(list(SH1_SCR_0.5mM = myReactome_ORAs[["SH1_0pt5mM-SCR_0pt5mM"]], SH2_SCR_0.5mM = myReactome_ORAs[["SH2_0pt5mM-SCR_0pt5mM"]],SCR_4mM_SCR_0.5mM =
myReactome_ORAs[["SCR_4mM-SCR_0pt5mM"]])) %>% dotplot(., showCategory=6) + ggtitle("Reactome ORA")+ scale_y_discrete(label=short_label)

```

```{r include=FALSE}
writeFile(myReactome_ORAs[[1]], paste0("myKEGG_ORAs_",names(myReactome_ORAs[1]),".csv"))
writeFile(myReactome_ORAs[[2]], paste0("myKEGG_ORAs_",names(myReactome_ORAs[2]),".csv"))
writeFile(myReactome_ORAs[[3]], paste0("myKEGG_ORAs_",names(myReactome_ORAs[3]),".csv"))
```


### Gene set enrichment analysis (gsea)

```{r cellular component gsea, eval=FALSE, message=FALSE, warning=FALSE, include=FALSE}
#### cellular component gsea
myCC_gsea = lapply(my_degs_rankingScoresList, function(x){
  gseGO(geneList = x, OrgDb = org.Mm.eg.db, ont = "CC", nPerm = 30000, keyType = "ENSEMBL",
        minGSSize = 10, maxGSSize = 500, pvalueCutoff = 0.05, pAdjustMethod = "BH",
        verbose = TRUE)})

dotplot(myCC_gsea[[1]], showCategory=6, split=".sign") + facet_grid(.~.sign)+ ggtitle(paste(names(myCC_gsea[1]),"Gene set enrichment analysis", sep = " ")) + scale_y_discrete(label=short_label)
dotplot(myCC_gsea[[2]], showCategory=6, split=".sign") + facet_grid(.~.sign)+ ggtitle(paste(names(myCC_gsea[2]),"Gene set enrichment analysis", sep = " ")) + scale_y_discrete(label=short_label)
dotplot(myCC_gsea[[3]], showCategory=6, split=".sign") + facet_grid(.~.sign)+ ggtitle(paste(names(myCC_gsea[3]),"Gene set enrichment analysis", sep = " ")) + scale_y_discrete(label=short_label)
```

#### biological process gsea
```{r gsea biological process, message=FALSE, warning=FALSE}
myBP_gsea = lapply(my_degs_rankingScoresList, function(x){
  gseGO(geneList = x, OrgDb = org.Mm.eg.db, ont = "BP", nPerm = 30000, keyType = "ENSEMBL",
        minGSSize = 10, maxGSSize = 500, pvalueCutoff = 0.05, pAdjustMethod = "BH",
        verbose = TRUE)})

dotplot(myBP_gsea[[1]], showCategory=6, split=".sign") + facet_grid(.~.sign)+ ggtitle(paste(names(myBP_gsea[1]),"Gene set enrichment analysis", sep = " ")) + scale_y_discrete(label=short_label)
dotplot(myBP_gsea[[2]], showCategory=6, split=".sign") + facet_grid(.~.sign)+ ggtitle(paste(names(myBP_gsea[2]),"Gene set enrichment analysis", sep = " ")) + scale_y_discrete(label=short_label)
dotplot(myBP_gsea[[3]], showCategory=6, split=".sign") + facet_grid(.~.sign)+ ggtitle(paste(names(myBP_gsea[3]),"Gene set enrichment analysis", sep = " ")) + scale_y_discrete(label=short_label)
```

```{r fig.width=9, fig.height=8}
normalized_data_SH1_SCR_0.5 = subset(normalized_data, select=c(gene, SCR_0pt5mM_A,SCR_0pt5mM_B,SH1_0pt5mM_A,SH1_0pt5mM_B))

serine1 = myBP_gsea[["SH1_0pt5mM-SCR_0pt5mM"]][which(myBP_gsea[["SH1_0pt5mM-SCR_0pt5mM"]]$ID=="GO:0009070"),]
serine1$core_enrichment = strsplit(as.character(serine1$core_enrichment), '\\/') 

SH1_SCR_0.5_serine_symbol = bitr(serine1$core_enrichment[[1]], fromType = "ENSEMBL", toType = "SYMBOL", OrgDb = org.Mm.eg.db) 

normalized_data_SH1_SCR_0.5_serine = normalized_data_SH1_SCR_0.5[(normalized_data_SH1_SCR_0.5$gene%in%
                                          SH1_SCR_0.5_serine_symbol$SYMBOL),]

my.colors <- colorRampPalette(rev(brewer.pal(n = 7, name ="RdYlBu")))(100)

drawheatmap = function(df, mytitle){
  mycounts = df
  rownames(mycounts)=mycounts$gene
  mycounts = subset(mycounts, select=-c(gene))
  pheatmap(mycounts, scale="row",cellwidth = 25,clustering_distance_cols = "euclidean", clustering_method = "complete", cellheight = 12, fontsize = 8, fontsize_row = 11, fontsize_col = 12, color = my.colors, border_color = "NA", cutree_cols=2, legend_breaks = c(-3,-2,-1,0,1,2,3), annotation_colors = anno_colors,angle_col = 45, main=mytitle)
}

SH1heatmap = drawheatmap(normalized_data_SH1_SCR_0.5_serine, "serine family amino acid biosynthetic process SH1_0pt5mM-SCR_0pt5mM")

#Running score and preranked list are traditional methods for visualizing GSEA result.
SH1gseaplot_BP = gseaplot2(myBP_gsea[["SH1_0pt5mM-SCR_0pt5mM"]], geneSetID = "GO:0009070", title = "serine family amino acid biosynthetic process SH1_0pt5mM-SCR_0pt5mM")
SH1gseaplot_BP
```

```{r include=FALSE}
writeFile(myBP_gsea[[1]], paste0("myBP_gsea_",names(myBP_gsea[1]),".csv"))
writeFile(myBP_gsea[[2]], paste0("myBP_gsea_",names(myBP_gsea[2]),".csv"))
writeFile(myBP_gsea[[3]], paste0("myBP_gsea_",names(myBP_gsea[3]),".csv"))
```

```{r molecular function gsea, eval=FALSE, message=FALSE, warning=FALSE, include=FALSE}
#### molecular function gsea
myMF_gsea = lapply(my_degs_rankingScoresList, function(x){
  gseGO(geneList = x, OrgDb = org.Mm.eg.db, ont = "MF", nPerm = 30000, keyType = "ENSEMBL",
        minGSSize = 10, maxGSSize = 500, pvalueCutoff = 0.05, pAdjustMethod = "BH",
        verbose = TRUE)})

dotplot(myMF_gsea[[1]], showCategory=6, split=".sign") + facet_grid(.~.sign)+ ggtitle(paste(names(myMF_gsea[1]),"Gene set enrichment analysis", sep = " ")) + scale_y_discrete(label=short_label)
dotplot(myMF_gsea[[2]], showCategory=6, split=".sign") + facet_grid(.~.sign)+ ggtitle(paste(names(myMF_gsea[2]),"Gene set enrichment analysis", sep = " ")) + scale_y_discrete(label=short_label)
dotplot(myMF_gsea[[3]], showCategory=6, split=".sign") + facet_grid(.~.sign)+ ggtitle(paste(names(myMF_gsea[3]),"Gene set enrichment analysis", sep = " ")) + scale_y_discrete(label=short_label)

```

#### kegg gsea
```{r kegg gsea, message=FALSE, warning=FALSE}
myKEGG_gsea <- lapply(my_degs_entrez_rankingScoresList, function(x){
  gseKEGG(geneList = x, organism = 'mmu', nPerm = 30000,minGSSize = 10, maxGSSize = 500, pvalueCutoff = 0.05, pAdjustMethod = "BH", verbose = TRUE)})

dotplot(myKEGG_gsea[[1]], showCategory=6, split=".sign") + facet_grid(.~.sign)+ ggtitle(paste(names(myKEGG_gsea[1]),"Gene set enrichment analysis", sep = " ")) + scale_y_discrete(label=short_label)
dotplot(myKEGG_gsea[[2]], showCategory=6, split=".sign") + facet_grid(.~.sign)+ ggtitle(paste(names(myKEGG_gsea[2]),"Gene set enrichment analysis", sep = " ")) + scale_y_discrete(label=short_label)
dotplot(myKEGG_gsea[[3]], showCategory=6, split=".sign") + facet_grid(.~.sign)+ ggtitle(paste(names(myKEGG_gsea[3]),"Gene set enrichment analysis", sep = " ")) + scale_y_discrete(label=short_label)

```

aminoacyl_tRNA_biosynthesis
```{r fig.width=10, fig.height=9}
normalized_data_SH1_SCR_0.5 = subset(normalized_data, select=c(gene, SCR_0pt5mM_A,SCR_0pt5mM_B,SH1_0pt5mM_A,SH1_0pt5mM_B))
normalized_data_SH2_SCR_0.5 = subset(normalized_data, select=c(gene, SCR_0pt5mM_A,SCR_0pt5mM_B,SH2_0pt5mM_A,SH2_0pt5mM_B))
aminoacyl_tRNA_biosynthesis1 = myKEGG_gsea[["SH1_0pt5mM-SCR_0pt5mM"]][which(myKEGG_gsea[["SH1_0pt5mM-SCR_0pt5mM"]]$ID=="mmu00970"),]
aminoacyl_tRNA_biosynthesis1$core_enrichment = strsplit(as.character(aminoacyl_tRNA_biosynthesis1$core_enrichment), '\\/') 

aminoacyl_tRNA_biosynthesis2 = myKEGG_gsea[["SH2_0pt5mM-SCR_0pt5mM"]][which(myKEGG_gsea[["SH2_0pt5mM-SCR_0pt5mM"]]$ID=="mmu00970"),]
aminoacyl_tRNA_biosynthesis2$core_enrichment = strsplit(as.character(aminoacyl_tRNA_biosynthesis2$core_enrichment), '\\/') 

# aminoacyl_tRNA_biosynthesis = union(aminoacyl_tRNA_biosynthesis1$core_enrichment[[1]],aminoacyl_tRNA_biosynthesis2$core_enrichment[[1]])

SH1_SCR_0.5_aminoacyl_tRNA_biosynthesis_symbol = bitr(aminoacyl_tRNA_biosynthesis1$core_enrichment[[1]], fromType = "ENTREZID", toType = "SYMBOL", OrgDb = org.Mm.eg.db) 

SH2_SCR_0.5_aminoacyl_tRNA_biosynthesis_symbol = bitr(aminoacyl_tRNA_biosynthesis2$core_enrichment[[1]], fromType = "ENTREZID", toType = "SYMBOL", OrgDb = org.Mm.eg.db) 

normalized_data_SH1_SCR_0.5_aminoacyl_tRNA_biosynthesis = normalized_data_SH1_SCR_0.5[(normalized_data_SH1_SCR_0.5$gene%in%
                                          SH1_SCR_0.5_aminoacyl_tRNA_biosynthesis_symbol$SYMBOL),]
normalized_data_SH2_SCR_0.5_aminoacyl_tRNA_biosynthesis = normalized_data_SH2_SCR_0.5[(normalized_data_SH2_SCR_0.5$gene%in%
SH2_SCR_0.5_aminoacyl_tRNA_biosynthesis_symbol$SYMBOL),]

my.colors <- colorRampPalette(rev(brewer.pal(n = 7, name ="RdYlBu")))(100)

drawheatmap = function(df, mytitle){
  mycounts = df
  rownames(mycounts)=mycounts$gene
  mycounts = subset(mycounts, select=-c(gene))
  pheatmap(mycounts, scale="row",cellwidth = 25,clustering_distance_cols = "euclidean", clustering_method = "complete", cellheight = 12, fontsize = 8, fontsize_row = 11, fontsize_col = 12, color = my.colors, border_color = "NA", cutree_cols=2, legend_breaks = c(-3,-2,-1,0,1,2,3), annotation_colors = anno_colors,angle_col = 45, main=mytitle)
}

#Running score and preranked list are traditional methods for visualizing GSEA result.
SH1gseaplot = gseaplot2(myKEGG_gsea[["SH1_0pt5mM-SCR_0pt5mM"]], geneSetID = "mmu00970", title = paste("Aminoacyl-tRNA biosynthesis", "SH1_0pt5mM-SCR_0pt5mM"))
SH2gseaplot = gseaplot2(myKEGG_gsea[["SH2_0pt5mM-SCR_0pt5mM"]], geneSetID = "mmu00970", title = paste("Aminoacyl-tRNA biosynthesis", "SH2_0pt5mM-SCR_0pt5mM"))

SH1heatmap = drawheatmap(normalized_data_SH1_SCR_0.5_aminoacyl_tRNA_biosynthesis, "Aminoacyl-tRNA biosynthesis SH1_0pt5mM-SCR_0pt5mM")
SH1gseaplot
SH2heatmap = drawheatmap(normalized_data_SH2_SCR_0.5_aminoacyl_tRNA_biosynthesis, "Aminoacyl-tRNA biosynthesis SH2_0pt5mM-SCR_0pt5mM")
SH2gseaplot
```

```{r include=FALSE}
writeFile(myKEGG_gsea[[1]], paste0("myKEGG_gsea_",names(myBP_gsea[1]),".csv"))
writeFile(myKEGG_gsea[[2]], paste0("myKEGG_gsea_",names(myBP_gsea[2]),".csv"))
writeFile(myKEGG_gsea[[3]], paste0("myKEGG_gsea_",names(myBP_gsea[3]),".csv"))

```


```{r Reactome gsea, eval=FALSE, message=FALSE, warning=FALSE, include=FALSE}
#### Reactome gsea
myReactome_gsea <- lapply(my_degs_entrez_rankingScoresList, function(x){
  gsePathway(geneList = x, organism = 'mouse', nPerm = 30000, minGSSize = 10, maxGSSize = 500, pvalueCutoff = 0.05, pAdjustMethod = "BH", verbose = TRUE)})

dotplot(myReactome_gsea[[1]], showCategory=6, split=".sign") + facet_grid(.~.sign)+ ggtitle(paste(names(myReactome_gsea[1]),"Gene set enrichment analysis", sep = " ")) + scale_y_discrete(label=short_label)
dotplot(myReactome_gsea[[2]], showCategory=6, split=".sign") + facet_grid(.~.sign)+ ggtitle(paste(names(myReactome_gsea[2]),"Gene set enrichment analysis", sep = " ")) + scale_y_discrete(label=short_label)
dotplot(myReactome_gsea[[3]], showCategory=6, split=".sign") + facet_grid(.~.sign)+ ggtitle(paste(names(myReactome_gsea[3]),"Gene set enrichment analysis", sep = " ")) + scale_y_discrete(label=short_label)

```

```{r save image}
save.image(file = "RNAseq.RData")
```
